Convergence rate of MAP estimates \\ for the exponential family

### 2021-09-30 main.tex
Move other idea to end of optimization part

\TODO{Tonight ! add equations for the full gaussian. I can dig them ou of my previous notebook.}

\TODO{add equations for other families, eg gaussian with mean and variance, or gamma or Dirichlet, or Categorical..}

@@

\subsection{Other Idea}
There are two main bottlenecks when casting MAP as an iterative algorithm. 
\begin{enumerate}
	\item The first one is the variance definition which always end up being infinite when asked to be uniform on every possible trajectories, if the log-partition  has some barrier structure.
	\item The second one is the step-size scheduling. $O(\inv{t})$ is decreasing too fast for most proof techniques. Even in the Euclidean setting where the variance does not matter, it results in a disappointing rate $O(\inv{\log t})$.
\end{enumerate}
Regarding this second point, it might be possible to do something comparable to \citet{gower2019sgd} or \citet{dragomir2021fast} who get a $O(\inv{t})$ rate with a $O(\inv{t})$ step-size. Take $m$ and $L$ as the (relative) strong-convexity and smoothness constants of the problem, along with $K=\frac{L}{m}$ the condition number.Their descent lemma looks like
\begin{align}
	B_{t+1} \leq ( 1 - \lr_t m) B_t + \lr_t^2 \sigma^2
\end{align}
where $B_t$ is some measure of suboptimality, in this case the $\ell^2$ distance between $\natp^t$ and $\natp^*$ ; and $\lr_t \leq \inv{2L}$.
On the other hand, the problem of most Bregman analysis \citep{hanzely2018fastest,dorazio2021stochastic} is that they do not get this $\lr^2$ in front the variance. Instead they simply get a $\lr$, because they do not benefit from the homogeneity of the norm
\begin{align}
	B_{t+1} \leq ( 1 - \lr_t m) B_t + \lr_t \sigma^2 \; .
\end{align}
The problem is that this exponent $2$ is crucial to get fast convergence with a $O(\inv{t})$ step-size. Otherwise the variance is too large and we need to decrease the step-size more slowly to reach the optimum quickly.
Now our question is what happens if we get something in-between ? e.g.
\begin{align}
	B_{t+1} \leq ( 1 - \lr_t m) B_t + \lr_t^\alpha \sigma^2 \; .
\end{align}
where $1<\alpha<2$ is some exponent characterizing the geometry of the problem.

@@

\subsection{Discussion}

\paragraph{Bound on the trajectory.}
Bound \eqref{eq:hypvariance} might not hold in general, but it can along the trajectory. Can we prove it using an inductive method such as \citet[Appendix A]{lacoste2012simpler}? In the case of the Gaussian variance, we cannot because $\mu$ takes every possible values as a random variable within the expectation.

@@

\begin{figure}[ht]
	\centering
	\includegraphics{figs/3pointslemma}
	\caption{Illustration of the other three points lemma -- eg the identity between Bregmans -- reproduced from \citet{eckstein1998approximate}.}
	\label{fig:3pointslemma}
\end{figure}

@@

\paragraph{Notations}
\begin{itemize}
    \item $\|\cdot\|_x = \sqrt{ \langle \nabla^2A^*(x)\cdot,\, \cdot \rangle }$
    \item $\|\cdot\|_x^*= \sqrt{ \langle [\nabla^2A^*(x)]^{-1}\cdot,\, \cdot \rangle }$
    \item $\lambda(x) = \|(A^*)'(x)\|_x^*$ (Newton decrements)
    \item $\omega = $
    \item $\omega^* = -t-\ln(1-t)$
\end{itemize}
Note that $\omega^*$ is convex and monotonne.

>>>>

\TODO{add equations for other families, eg gamma or Dirichlet, or Categorical..}

@@

@@

\section{Discussion and other idea}
There are two main bottlenecks when casting MAP as an iterative algorithm. 
\begin{enumerate}
	\item The first one is the variance definition which always end up being infinite when asked to be uniform on every possible trajectories, if the log-partition  has some barrier structure.
	\item The second one is the step-size scheduling. $O(\inv{t})$ is decreasing too fast for most proof techniques. Even in the Euclidean setting where the variance does not matter, it results in a disappointing rate $O(\inv{\log t})$.
\end{enumerate}
Regarding this second point, it might be possible to do something comparable to \citet{gower2019sgd} or \citet{dragomir2021fast} who get a $O(\inv{t})$ rate with a $O(\inv{t})$ step-size. Take $m$ and $L$ as the (relative) strong-convexity and smoothness constants of the problem, along with $K=\frac{L}{m}$ the condition number.Their descent lemma looks like
\begin{align}
	B_{t+1} \leq ( 1 - \lr_t m) B_t + \lr_t^2 \sigma^2
\end{align}
where $B_t$ is some measure of suboptimality, in this case the $\ell^2$ distance between $\natp^t$ and $\natp^*$ ; and $\lr_t \leq \inv{2L}$.
On the other hand, the problem of most Bregman analysis \citep{hanzely2018fastest,dorazio2021stochastic} is that they do not get this $\lr^2$ in front the variance. Instead they simply get a $\lr$, because they do not benefit from the homogeneity of the norm
\begin{align}
	B_{t+1} \leq ( 1 - \lr_t m) B_t + \lr_t \sigma^2 \; .
\end{align}
The problem is that this exponent $2$ is crucial to get fast convergence with a $O(\inv{t})$ step-size. Otherwise the variance is too large and we need to decrease the step-size more slowly to reach the optimum quickly.
Now our question is what happens if we get something in-between ? e.g.
\begin{align}
	B_{t+1} \leq ( 1 - \lr_t m) B_t + \lr_t^\alpha \sigma^2 \; .
\end{align}
where $1<\alpha<2$ is some exponent characterizing the geometry of the problem.

@@

\begin{figure}[ht]
	\centering
	\includegraphics[width=.7\textwidth]{figs/3pointslemma}
	\caption{Illustration of the other three points lemma -- eg the identity between Bregmans -- taken from \citet{eckstein1998approximate}.}
	\label{fig:3pointslemma}
\end{figure}

@@

\paragraph{Notations}
\begin{itemize}
    \item $\|\cdot\|_x = \sqrt{ \langle \nabla^2A^*(x)\cdot,\, \cdot \rangle }$
    \item $\|\cdot\|_x^*= \sqrt{ \langle [\nabla^2A^*(x)]^{-1}\cdot,\, \cdot \rangle }$
    \item $\lambda(x) = \|(A^*)'(x)\|_x^*$ (Newton decrements)
    \item $\omega = t-\ln (1+t)$
    \item $\omega^* = -t-\ln(1-t)$
\end{itemize}
Note that $\omega^*$ is convex and monotonne.
